The Atala PRISM Protocol

### 2022-12-20 prism-backend/research/prism-protocol/did_atala.tex
Proving UC security of Atala DIDs.

At a high level, to apply the results in \cite{bch+20,bzh21}, we show the global
subroutines used in \IdealGPKIDID and \RealPKIDIDAtala can be merged into the
single ``meta'' global subroutine $\hat{\gamma}_{\gamma_1,\dots,\gamma_n}$
required by \thmref{thm:uc-bhz21}, which furthermore meets the conditions to be
subsequently suitable for UC-emulation. In addition, we argue that this is
usually the case for the known realisations of each respective global
subroutine. (This ensures that our approach is natural, and indeed UC-realisable
in practice.)
%
We then describe how the management protocol \UCM works in conjunction with
our \IdealGPKIDID and the ``meta'' global subroutines (in the ideal world), and
with \RealPKIDIDAtala and the ``meta'' global subroutine (in the real world).
%
Finally, we show that the latter emulates the former.
%
We emphasise that this is merely an application of the results in
\cite{bch+20,bhz21} to our setting. The previously sketched process is detailed
next, directly referencing the relevant theorems of the relevant works, when
appropriate.

\paragraph{The meta-global subroutine \gAtala.} %
We next show that the global subroutines used in \RealPKIDIDAtala are leveraged
in a way that is compatible with \cite[Theorem 4.3]{bhz21}. This is not a full
proof for a concrete application of such theorem, as that would require proving
that concrete protocols UC-emulate each global functionality. However, we
initially leave it ``incomplete'' to show that any combination of instantiations
of our global subroutines can be used to build a secure variant of the Atala
DID protocol (later, we show the concrete instantiations that Atala currently
uses). For concreteness, we refer to this ``partial'' property as mergability
of the global subroutines of Atala into \gAtala.

\begin{lemma}[Mergability of \gAtala]
  \label{lma:gAtala}
  Let $\gamma_1 \gets \IdealGdledger$, $\gamma_2 \gets \IdealGRO$, $\gamma_3
  \gets \IdealGclock$. Then:
  \begin{enumerate}
  \item For $i < 3$, $\gamma_i$ is subroutine respecting except for calls to
    $\gamma_{i+1,\dots,3}$, and $\gamma_3$ is subroutine respecing. All
  $\gamma_i$'s are subroutine exposing.
\item For $i>1$, $\gamma_i$ is $\gamma_j$-regular, for all $j \in \set{1,
    \dots,\i-1}$.
  \end{enumerate}
\end{lemma}

\begin{proof}
  \qed
\end{proof}

>>>>

At a high level, to apply the results in \cite{bch+20,bhz21}, we show the global
subroutines used in \IdealGPKIDID and \RealPKIDIDAtala can be merged into the
single ``meta'' global subroutine $\hat{\gamma}_{\gamma_1,\dots,\gamma_n}$
required by \thmref{thm:uc-bhz21}, which furthermore meets the conditions to be
subsequently suitable for UC-emulation. In addition, we argue that this is
usually the case for the known realisations of each respective global
subroutine. (This suggests that our approach is natural, and indeed
UC-realisable in practice.)
%
We then describe how the management protocol \UCM works in conjunction with
our \IdealGPKIDID and the ``meta'' global subroutines in the ideal world (i.e.,
$\UCM[\IdealGPKIDID,\hat{\gamma}_{\gamma_1,\dots,\gamma_n}]$), and
with \RealPKIDIDAtala and the ``meta'' global subroutine in the real world
(i.e., $\UCM[\RealPKIDIDAtala,\hat{\gamma}_{\gamma_1,\dots,\gamma_n}]$).
%
Finally, we show that the latter emulates the former. This concludes the proof
for \thmref{thm:uc-bhz21}.
%
We emphasise that this is merely an application of the results in
\cite{bch+20,bhz21} to our setting. The previously sketched process is detailed
next, directly referencing the relevant theorems of the corresponding works when
appropriate.

\paragraph{The meta-global subroutine \gAtala.} %
\todo{The idea is here, but the wording is clumsy. Re-write.}
We next show that the global subroutines used in \RealPKIDIDAtala are leveraged
in a way that is compatible with \cite[Theorem 4.3]{bhz21}. This is not a full
proof for a concrete application of the theorem, as that would require proving
that concrete protocols UC-emulate each global functionality. However, we
initially leave it ``incomplete'' to show that any combination of instantiations
of our global subroutines can be used to build a secure variant of the Atala
DID protocol (later, we show the concrete instantiations that Atala currently
uses). For concreteness, we refer to this ``partial'' property as mergability
of the global subroutines of Atala into \gAtala.

\begin{lemma}[Mergeability of \gAtala]
  \label{lma:gAtala}

  Let $\gamma_1 \gets \IdealGdledger$, $\gamma_2 \gets \IdealGRO$, $\gamma_3
  \gets \IdealGclock$. Then:

  \begin{enumerate}
  \item For $i < 3$, $\gamma_i$ is subroutine respecting except for calls to
    $\gamma_{i+1,\dots,3}$, and $\gamma_3$ is subroutine respecing. All
    $\gamma_i$'s are subroutine exposing.
  \item For $i>1$, $\gamma_i$ is $\gamma_j$-regular, for all $j \in \set{1,
      \dots,i-1}$.
  \end{enumerate}

\end{lemma}

\begin{proof}
  \todo{In the sequel, when the $\gamma_i$'s should be protocols, not ideal
    functionalities; e.g., $\gamma_3 = \IdealP_{\IdealGclock}$, not
    $\IdealGclock$ (the argument should be ok, it is just a wording error).}
  We start with the first item, for which we briefly recall the notion of
  subroutine respecting protocols. Intuitively, a protocol $\pi$ is subroutine
  respecting if, for any session \sid of $\pi$, the main parties of \sid are the
  only ones that communicate with the outside world; this also means that
  subparties only accept input and subroutine output from parties in \sid. A
  more precise definition is available in the 2020 version of \cite{canetti00}.

  Now, the protocol $\IdealP_\IdealF$ associated to an ideal functionality
  \IdealF is composed of dummy parties that simply relay information back and
  forth between the environment \Env and \IdealF (and no one else). The dummy
  parties are the main parties of $\IdealP_\IdealF$, whereas \IdealF is a
  sub-party.

  In the case of \IdealGclock ($\gamma_3$) and \IdealGRO ($\gamma_2$), their
  corresponding protocols, $\IdealP_{\IdealGclock}$ and $\IdealP_{\IdealGRO}$
  are trivially subroutine respecting. Namely, for each session \sid of
  $\IdealP_{\IdealGclock}$ (resp. $\IdealP_{\IdealGRO}$):

  \begin{itemize}
  \item \IdealGclock (resp. \IdealGRO) is the only sub-party in the protocol
    session \sid, and it only accepts input from the main (dummy) parties in
    $\IdealP_{\IdealGclock}$ (resp. $\IdealP_{\IdealGRO}$) in \sid.
  \item The main (dummy) parties of session \sid in $\IdealP_{\IdealGclock}$
    (resp. $\IdealP_{\IdealGRO}$) ignore subroutine outputs from parties other
    than \IdealGclock (resp. \IdealGRO)'s on session \sid. Similarly,
    \IdealGclock (resp. \IdealGRO) ignores subroutine outputs.
  \item \IdealGclock (resp. \IdealGRO) in \sid only passes subroutine output to
    the main (dummy) parties in \sid.
  \item Main (dummy) parties in session \sid of $\IdealP_{\IdealGclock}$ (resp.
    $\IdealGRO$), and \IdealGclock (resp. \IdealGRO)  don't pass input to
    parties not in \sid. 
  \end{itemize}

  Next, observe that \IdealGdledger ($\gamma_1$) is also subroutine respecting
  except for calls to \IdealGclock ($\gamma_3$). Concretely, for each session
  of the ideal protocol $\IdealP_{\IdealGdledger}$ of \IdealGdledger:

  \begin{itemize}
  \item Excluding calls to \IdealGclock, \IdealGdledger is the only sub-party in
    the protocol session \sid, and it only accepts input from the main (dummy)
    parties in $\IdealP_{\IdealGdledger}$ in \sid.
  \item The main (dummy) parties of session \sid in $\IdealP_{\IdealGdledger}$
    ignore subroutine outputs from parties other than \IdealGdledger's on
    session \sid. Also, \IdealGdledger ignores subroutine outputs from any
    party, with the exception from parties in \IdealGclock.
  \item \IdealGdledger in \sid only passes subroutine output to the main (dummy)
    parties of $\IdealP_{\IdealGdledger}$, and to the parties in \IdealGclock,
    in \sid.
  \item Main (dummy) parties in session \sid of $\IdealP_{\IdealGdledger}$ and
    \IdealGdledger don't pass input to parties not in \sid, with the exception
    of parties in $\IdealP_{\IdealGclock}$.
  \end{itemize}

  In summary, $\gamma_3 = \IdealGclock$ and $\gamma_2 = \IdealGRO$ are
  subroutine respecting (and, obviously, $\gamma_2$ is also subroutine
  respecting except for calls to $\gamma_3$). Also, $\gamma_1 = \IdealGdledger$
  is subroutine respecting except for calls to $\gamma_3$ -- and, since
  $\gamma_1$ does not use $\gamma_2$, it is trivially also subroutine respecting
  except for calls to $\gamma_2$ and $\gamma_3$.

  Thus, condition 1 of \lmaref{lma:gAtala} is met. Now, we move to condition 2.

  \qed
\end{proof}

\paragraph{The management protocol $\UCM[\IdealGPKIDID,\gAtala]$.}

\paragraph{The management protocol $\UCM[\RealPKIDIDAtala,\gAtala]$.}

\paragraph{$\UCM[\IdealGPKIDID,\gAtala]$ $\xi$-UC emulates
  $\UCM[\RealPKIDIDAtala,\gAtala]$.}

### 2022-12-20 prism-backend/research/prism-protocol/preliminaries.tex
Proving UC security of Atala DIDs.

    \centering
    \procedure{$\pcif \cmd = \uccmd{HashQuery} \land \arg = \msg$}{%
      \pcif (\msg,h) \notin \hashlist:
      h \getr \bin^{l}; \hashlist \gets \hashlist \cup \lbrace (\msg,h) \rbrace \\
      \ucio{Output}~(\uccmd{HashConfirm},h)
    }
  \end{framed}
\end{figure}

@@

\begin{figure}[ht]
  \begin{framed}
    \begin{minipage}[t]{\textwidth}
      \textrm{Upon receiving a call $\IdealGclock(\cmd,\arg)$ from
        some party $P$ to run command \cmd, with arguments \arg,
        \IdealGclock operates as follows.} \\
      \textrm{\IdealGclock maintains a variable $t$ with the current itme,
        a variable $T$ with the list of parties keeping track of the time,
        and a map $A$ which keeps the parties' (dis)agreements to advance the
        clock.}
    \end{minipage}
    \vspace*{0.5em}

    \begin{minipage}[t]{0.40\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Register} \land \arg = \emptyset$}{
        T \gets T \cup \set{P} \\
        \ucio{Output}~(\uccmd{Registered})~\text{to}~P \\
      }
      \procedure{$\pcif \cmd = \uccmd{Update} \land \arg = \emptyset$}{
        A(P) \gets \top \\
        \pcif \forall P \in T, A(P) = \top: \\
        \pcind t \gets t + 1; \forall P \in T, A(P) = \bot \\
        \pcind \ucio{Send}~(\uccmd{Updated})~\text{to}~P        
      }
    \end{minipage}
    \hspace*{1.25em}
    \begin{minipage}[t]{0.60\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Deregister} \land \arg = \emptyset$}{
        T \gets T \setminus \set{P} \\
        \text{Output}~(\uccmd{DeregisteredOk})~\text{to}~P \\
      }
      \procedure{$\pcif \cmd = \uccmd{Read} \land \arg = \emptyset$}{
        \text{Output}~(\uccmd{ReadOk},t)~\text{to}~P
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal subroutine for a global clock, from \cite{kkk21}.}
  \label{fig:gclock}
\end{figure}

@@

\begin{figure}[ht]
  \begin{framed}
    \begin{minipage}[t]{\textwidth}
      \textrm{Upon receiving a call $\IdealGdledger(\cmd,\arg)$ from
        some party $P$ to run command \cmd, with arguments \arg,
        \IdealGdledger operates as follows.} \\
      \textrm{\IdealGdledger maintains a mapping \ldgMap which associates
        parties with their view of the ledger.
        \ldgUtxo is a multiset of unconfirmed transactions.
        \ldgState is the authoritative ledger state
        All mappings and (multi-)sets are initialised to empty values.
      }
    \end{minipage}
    \vspace*{0.5em}

    \begin{minipage}[t]{0.40\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Submit} \land \arg = \ldgTx$}{
        \ucio{Send}~(\uccmd{Read})~\text{to}~\IdealGclock \\
        \ucio{Recv}~(\uccmd{ReadOk},t)~\text{from}~\IdealGclock \\
        \ldgUtxo \gets \ldgUtxo \cup \set{(\ldgTx,t)} \\ 
        \ucio{Output}~(\uccmd{Transaction},\ldgTx,t)~\text{to}~\adv \\
      }
      \procedure{$\pcif \cmd = \uccmd{Extend} \land \arg = \ldgState'$}{
        \pcif P = \adv \land \ldgState' \subseteq \ldgUtxo: \\
        \pcind \ldgUtxo \gets \ldgUtxo \setminus \ldgState' \\
        \pcind \ldgState \gets \ldgState || \ldgState'
      }
    \end{minipage}
    \hspace*{1.25em}
    \begin{minipage}[t]{0.60\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Read} \land \arg =
        \emptyset$}{
        % \pcif P = \adv: data \gets \ldgState \\
        % \pcelse: data \gets \ldgMap(P) \\
        \ucio{Send}~(\uccmd{Read})~\text{to}~\IdealGclock \\
        \ucio{Recv}~(\uccmd{ReadOk},t)~\text{from}~\IdealGclock \\
        \pcif \exists (\ldgTx,t') \in \ldgUtxo~\suchthat~|t-t'| > \delta:
        \pcreturn \bot \\
        \pcelse \pcif \exists (\ldgTx,t') \in \ldgState~\suchthat~ |t-t'|
        > \delta~\land \\
        \pcind \exists P' \in \ldgHonest~\suchthat~(\ldgTx,t') \notin \ldgMap(P'):
        \pcreturn \bot \\        
        \text{Output}~(\uccmd{ReadOk}, \ldgHead(\ldgMap(P)))~\text{to}~P \\
      }
      \procedure{$\pcif \cmd = \uccmd{Advance} \land \arg = (P',\ldgState')$}{
        \pcif P = \adv \land \ldgMap(P') \prec \ldgState' \prec \ldgState:
        \ldgMap(P') \gets \ldgState'
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal subroutine for a global $\delta$-delay ledger, from
    \cite{kkk21}.}
  \label{fig:gledger}
\end{figure}

>>>>

    \centering
    \procedure{$\pcif \cmd = \uccmd{HashQuery} \land \arg = \msg$}{%
      \pcif (\msg,h) \notin \hashlist:
      h \getr \bin^{l}; \hashlist \gets \hashlist \cup \lbrace (\msg,h) \rbrace \\
      \ucio{Output}~(\uccmd{HashConfirm},\sid,h)
    }
  \end{framed}
\end{figure}

@@

\begin{figure}[ht]
  \begin{framed}
    \begin{minipage}[t]{\textwidth}
      \textrm{Upon receiving a call $\IdealGclock(\cmd,\sid,\arg)$ from
        some party $P$ to run command \cmd, in session \sid, with arguments
        \arg, \IdealGclock operates as follows.} \\
      \textrm{\IdealGclock maintains a variable $t$ with the current time,
        a variable $T$ with the list of parties keeping track of the time,
        and a map $A$ which keeps the parties' (dis)agreements to advance the
        clock.}
    \end{minipage}
    \vspace*{0.5em}

    \begin{minipage}[t]{0.40\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Register} \land \arg = \emptyset$}{
        T \gets T \cup \set{P} \\
        \ucio{Output}~(\uccmd{Registered},\sid)~\text{to}~P \\
      }
      \procedure{$\pcif \cmd = \uccmd{Update} \land \arg = \emptyset$}{
        A(P) \gets \top \\
        \pcif \forall P \in T, A(P) = \top: \\
        \pcind t \gets t + 1; \forall P \in T, A(P) = \bot \\
        \pcind \ucio{Send}~(\uccmd{Updated},\sid)~\text{to}~P        
      }
    \end{minipage}
    \hspace*{1.25em}
    \begin{minipage}[t]{0.60\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Deregister} \land \arg = \emptyset$}{
        T \gets T \setminus \set{P} \\
        \text{Output}~(\uccmd{DeregisteredOk},\sid)~\text{to}~P \\
      }
      \procedure{$\pcif \cmd = \uccmd{Read} \land \arg = \emptyset$}{
        \text{Output}~(\uccmd{ReadOk},\sid,t)~\text{to}~P
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal subroutine for a global clock, from \cite{kkk21}.}
  \label{fig:gclock}
\end{figure}

@@

\begin{figure}[ht]
  \begin{framed}
    \begin{minipage}[t]{\textwidth}
      \textrm{Upon receiving a call $\IdealGdledger(\cmd,\sid,\arg)$ from
        some party $P$ to run command \cmd, in session \sid, with arguments
        \arg, \IdealGdledger operates as follows.} \\
      \textrm{\IdealGdledger maintains a mapping \ldgMap which associates
        parties with their view of the ledger.
        \ldgUtxo is a multiset of unconfirmed transactions.
        \ldgState is the authoritative ledger state
        All mappings and (multi-)sets are initialised to empty values.
      }
    \end{minipage}
    \vspace*{0.5em}

    \begin{minipage}[t]{0.40\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Submit} \land \arg = \ldgTx$}{
        \ucio{Send}~(\uccmd{Read},\sid)~\text{to}~\IdealGclock \\
        \ucio{Recv}~(\uccmd{ReadOk},\sid,t)~\text{from}~\IdealGclock \\
        \ldgUtxo \gets \ldgUtxo \cup \set{(\ldgTx,t)} \\ 
        \ucio{Output}~(\uccmd{Submitted},\sid,\ldgTx,t)~\text{to}~\adv \\
      }
      \procedure{$\pcif \cmd = \uccmd{Extend} \land \arg = \ldgState'$}{
        \pcif P = \adv \land \ldgState' \subseteq \ldgUtxo: \\
        \pcind \ldgUtxo \gets \ldgUtxo \setminus \ldgState' \\
        \pcind \ldgState \gets \ldgState || \ldgState'
      }
    \end{minipage}
    \hspace*{1.25em}
    \begin{minipage}[t]{0.60\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Read} \land \arg =
        \emptyset$}{
        % \pcif P = \adv: data \gets \ldgState \\
        % \pcelse: data \gets \ldgMap(P) \\
        \ucio{Send}~(\uccmd{Read},\sid)~\text{to}~\IdealGclock \\
        \ucio{Recv}~(\uccmd{ReadOk},\sid,t)~\text{from}~\IdealGclock \\
        \pcif \exists (\ldgTx,t') \in \ldgUtxo~\suchthat~|t-t'| > \delta:
        \pcreturn \bot \\
        \pcelse \pcif \exists (\ldgTx,t') \in \ldgState~\suchthat~ |t-t'|
        > \delta~\land \\
        \pcind \exists P' \in \ldgHonest~\suchthat~(\ldgTx,t') \notin \ldgMap(P'):
        \pcreturn \bot \\        
        \text{Output}~(\uccmd{ReadOk},\sid,\ldgHead(\ldgMap(P)))~\text{to}~P \\
      }
      \procedure{$\pcif \cmd = \uccmd{Advance} \land \arg = (P',\ldgState')$}{
        \pcif P = \adv \land \ldgMap(P') \prec \ldgState' \prec \ldgState:
        \ldgMap(P') \gets \ldgState'
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal subroutine for a global $\delta$-delay ledger, from
    \cite{kkk21}.}
  \label{fig:gledger}
\end{figure}
